\documentclass[12pt]{letter}

% \usepackage[britdate]{SU-letter}
\usepackage[britdate]{UH-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[round]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

%\usepackage{natbib}
% \bibpunct[; ]{(}{)}{;}{a}{,}{;}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{Postdoctoral fellow}
\department{Department of Agricultural Sciences}
\location{University of Helsinki}
\cityzip{Helsinki, Finland}
\telephone{}
\fax{}
\email{alyssa.cirtwill@helsinki.fi}
\url{http://cirtwill.github.io}
\name{Dr. Alyssa R. Cirtwill}

\begin{document}

\begin{letter}{\bf Prof. Dries Bonte [[[CHECK SPEC CHARS]]]\\
Oikos Editorial Office \\
Lund University \\
Lund, Sweden}


\opening{Dear Prof. Bonte:}

Thank you for the time and effort you have put into reviewing our manuscript, and for the helpful feedback provided. In light of this feedback, we have made a number of changes to the manuscript. We believe these changes have substantially strengthened the manuscript. In particular, we have substantially simplified the methods and this section should now be much clearer. We now also look at \emph{where} in a motif species are (their motif position) as opposed to simply which motifs they appear in, in order to delve more deeply into how this may affect their stability. 

We must also apologise for the long delay in returning this manuscript. It has been a particularly busy and difficult time for a number of reasons (largely Covid related), including one author moving continents. We also have to confess to having missed the statement of the deadline, so did not ask for an extension of time. We hope you will still accept this as a resubmission. 

Below is a point-by-point response to the reviewers comments. Our responses follow the reviewers comments and are preceded by a bold \textbf{R}. Sections quoted from the manuscript are indented.

%   My co-author and I are happy to invite you to consider our manuscript
%   \emph{Stable motifs delay species loss in simulated food webs} for publication in \textbf{Oikos}.


%   Motifs --unique configurations of small numbers of interacting species-- are becoming more common as a tool for describing the meso-scale structure of ecological networks (i.e., a species 'neighborhood' within the network, including direct and close indirect interactions). 
%   The set of motifs a species participates in has already been linked to species' traits, taxonomy, and location.
%   More recent research has also suggested that some 'stable' motifs, in isolation, dampen disturbances and facilitate persistence of the species within them while others do not.
%   This suggests that species which participate in many of these `stable' motifs should be more likely to persist following a disturbance than species which participate in more of the `unstable' motifs.
%   If this is the case, motifs offer the possibility of estimating extinction risk within all types of ecological communities, without necessarily having to know the structure of the entire network.


%   Here we test this possibility using a large set of realistic simulated networks. In brief, we find that participating in more of the stable motifs is generally associated with longer persistence. Participating more often in the omnivory motif, however, is associated with shorter persistence. Motifs can therefore be added to the toolbox of researchers wishing to estimate species' risk of extinction. In particular, motifs offer a more detailed perspective on network structure than simpler measures such as degree (number of interaction partners) or trophic level and may be useful when comparing the risk of apparently similar taxa.


%   Our results also offer a potential explanation for the variable effects of omnivory on community stability. In our networks, participating more often in the omnivory motif was associated with greater extinction risk. However, species with high numbers of interaction partners (which tend to have lower extinction risk) appeared more often in the omnivory motif than species with few interaction partners. The apparent effect of omnivory may therefore depend on the balance between the effect of omnivory \emph{per se} and the effect of numbers of interaction partners.


%   Predicting which species are most likely to go extinct following a disturbance is one of the pressing ecological tasks of our time. 
%   Motifs can contribute to solving this problem as they provide a rich description of the ways species fit into their communities which can be applied to any ecological community and, unlike whole-network measures of stability, can be calculated for a focal species or group even if the structure of an entire community is unknown.
%   Moreover, our results advance the basic science of network ecology by establishing the relationships between motif participation, degree, and trophic level. 
%   Because of the broad applicability and combination of theoretical and practical relevance of our results, we believe that \textbf{Oikos} is the best place to publish this manuscript. 
%   We hope that you will agree and eagerly await your reply.


%   A non-peer reviewed preprint of this manuscript has been posted to bioRxiv \\\noindent(doi: https://www.biorxiv.org/content/10.1101/2021.04.06.438635).

% \closing{Best regards,}

01-Jul-2021

Dear Dr. Cirtwill:

Manuscript ID OIK-08610 entitled ``Stable motifs delay species loss in simulated food webs'' which you submitted to \emph{Oikos}, has been reviewed.  The comments of the reviewer(s) and the recommendation by the SE are included at the bottom of this letter.

Based on the referee(s) comments, the subject editor has recommended substantial revisions to your manuscript.  Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript.

At this stage, we strongly encourage authors to contribute with text (maximum 200 words each) for the sessions Speculations and Alternative Viewpoints (among co-authors) to appear after the Discussion session.  These sessions are not compulsory and aim at increasing the visibility of your published work and encourage debate in ecology. Contributions to these sessions can be after the manuscript undergoes peer review or just in the version before final acceptance for publication.   For detailed information go to: http://www.oikosjournal.org/authors/author-guidelines

Author Contribution Indication
The contributions of each author to this work must now be indicated when you submit your revised manuscript. To add Author Contributions using CRediT taxonomy (http://credit.niso.org/contributor-roles-defined/), simply click the ``Provide CRediT Contribution'' link for each author in the ‘Authors & Institutions’ step of the submission process. From there, you will be able to check applicable Author/Contributor Roles and, if available, specify the Degree of Contribution. You MUST provide this information as part of the revision process. Author Contributions will be published with the accepted article and cannot be edited after article acceptance. Therefore you must ensure the Author Contribution information you provide is accurate prior to final acceptance.

There are two ways to submit your revised manuscript. You may use the link below to submit your revision online with no need to enter log in details:

*** PLEASE NOTE: This is a two-step process. After clicking on the link, you will be directed to a webpage to confirm. ***

https://mc.manuscriptcentral.com/oikos?URL_MASK=745eb08638f94600bc8791bf54bdd55a

Alternatively log into https://mc.manuscriptcentral.com/oikos and enter your Author Center. You can use the revision link or you will find your manuscript title listed under "Manuscripts with Decisions."  Under "Actions," click on "Create a Revision."  Your manuscript number has been appended to denote a revision.  Please DO NOT upload your revised manuscripts as a new submission.

You will be unable to make your revisions directly on the original uploaded version of the manuscript.  Instead, revise your manuscript using a word processing program and save it on your computer.

Please highlight the changes to your manuscript within the document by using the track changes mode in MS Word or by using bold or colored text. Once the revised manuscript is prepared, you can upload it and submit it through your Author Center.

Please revise the manuscript carefully in line with suggestions made by the reviewer(s) and respond to the criticism in the space provided or as a separate uploaded cover letter.  In order to expedite the processing of the revised manuscript, please be as specific as possible in your response to the reviewer(s). Please, upload a file with your response to the reviewers' comments when submitting your revised version (make sure that the file is also anonymous)

IMPORTANT:  Your original files are available to you when you upload your revised manuscript.  Please delete any redundant files before completing the submission.

TEXT: Please read the instruction for authors thoroughly before resubmitting your manuscript.

FIGURES: Please ensure that your illustrations are of high quality (> 300 dpi), preferably as high resolution jpeg, pdf or gifs. Please avoid pixelated symbols and legends. The normal width of an illustration in Oikos is 7.5 cm which means that all legends and symbols should be easily readable at this size. When scientifically motivated, a larger width, 11 cm, is also available.

REFERENCES: Please revise the reference list so that it complies with our style as seen in current issues of Oikos.  Papers with incomplete or malformatted reference lists, will not be officially accepted for publication.

MANUSCRIPT DETAILS: If any changes have been made to the title and/or abstract, please ensure that the relevant parts in the manuscript details within your Author Center are updated accordingly.

Because we are trying to facilitate timely publication of manuscripts submitted to Oikos, your revised manuscript should be uploaded as soon as possible and not later than within 1 month from today.  If it is not possible for you to submit your revision in a reasonable amount of time, we may have to consider your paper as a new submission.  If you feel that you will be unable to submit your revision within the time allowed please contact me to discuss the possibility of extending the revision time.

PLEASE: Submit your revised manuscript as MS Word or OpenOffice file(s). Submit your manuscript as one complete file, which includes: the text, legends, illustrations, tables and references.

Also submit, as separate files, all illustrations in high resolution (see above).

Once again, thank you for submitting your manuscript to Oikos and I look forward to receiving your revision.

Wiley Editing Services Available to All Authors
Should you be interested, Wiley Editing Services offers expert help with manuscript, language, and format editing, along with other article preparation services. You can learn more about this service option at www.wileyauthors.com/eeo/preparation. You can also check out Wiley’s collection of free article preparation resources for general guidance about writing and preparing your manuscript at www.wileyauthors.com/eeo/prepresources.

Sincerely,

Prof. Dries Bonte
Editor-in-Chief, Oikos




\clearpage

\Large{\textbf{Reply to Subject Editor}}

    
    \begin{quotation}    
  Recommendation by the Subject Editor (Dr. François Munoz):

  The reviewers and me agree that the topic of the paper is interesting and fits well the scope of Oikos.
  Understanding how the participation of species to trophic network motifs relates to species extinction risk should convey novel theoretical insights on the dynamics of trophic interaction networks.

  As Reviewer 2 noticed, I found that the methodology is quite complex, regarding the simulation of networks and their dynamics, and also regarding the statistical analyses of the link between network-based metrics and species extinction risk.
  Therefore, I recommend that the authors carefully revise the manuscript in the light of the comments and suggestions from both reviewers. The revision works should allow simplifying the presentation of results and thereby make them easier to understand.

  I look forward receiving a revised version of your work.

  Sincerely,

  François
  \end{quotation}
  
  [Alyssa]
  We have taken to heart the point which you and Reviewer 2 have raised about the complexity of our methodology. We have substantially revised our statistical analyses to better focus on our key questions: 1) are species' motif roles related to their risk of extinction? and 2) if so, which motifs show the strongest relationships? We have also revised the text to more clearly explain our simulation methodology. We hope that this revision is clearer and easier to follow.

\clearpage

\Large{\textbf{Reply to Reviewer 1}} [still needs preface.]

\textbf{R:} Reviewer 1's only major wish was that we make the code available, which we now do (see below). We also address their minor comments below. Many of these are no longer relevant due to our simplification of methods. Those that are relevant have helped us refine and clarify the manuscript.

\textbf{Comments by Reviewer 1:}
  In this work the author present a modelling study to explore how a species' participation in network motifs that are either stable or unstable relates to their extinction risk following a perturbation. The authors also discuss how species roles relate to other common network properties like species' degree and trophic position. 

  My only major wish is that the code was made available either as a supplement, or in a repository such as Zenodo or Figshare. The methods were fairly clear, but being able to review the code would be very helpful.


  \textbf{R:}  
  We now make our code available as a supplement.


  \textbf{Minor comments:}

  1. Resolution of time to extinction [done]


    Why is mean time to extinction computed by simulating multiple rounds of 10 timesteps rather than simulating the 500 timesteps and identifying where extinctions occur? Does this mean that time to extinction is defined as number of rounds until extinction, or by counting the timesteps within each round until the biomass falls below 10^-5? 

    \textbf{R:} This was done due to computational limits when running the models and storing output. We felt that a robust sample size and reasonably broad range of network structures were more important than obtaining very precise measures of time to extinction (especially since we do not simulate any particular species of interest). Time to extinction is defined as number of rounds until extinction - we have revised the text to make this clearer. The revised section now reads:

    \begin{quotation}

            Once we identified species' roles in the equilibrium networks, we perturbed the networks by removing a single species. 
            We then simulated post-removal community dynamics for 500 time-steps.
            After every 10 time-steps, any species with a biomass below our threshold of 1$\times10^{-5}$ was considered to have gone extinct and its biomass was set to 0.
            This reduced temporal resolution was chosen due to computational limitations and allowed us to maintain a large set of simulated networks.
            For each species other than the artificially removed species, we defined the time to extinction as the round of 10 time-steps in which the species' biomass fell below our threshold. 

    \end{quotation}


    2. Role dispersion not clear [done]


        I am not clear on what role dispersion (line 146) is in this case. Is this the variance of the counts for each motif per species? In other words, what does it mean to have a more or less variable role (line 208)?
        
        \textbf{R:} Dispersion is, essentially, a multivariate extension of variance which takes counts of all motifs into account simultaneously. 
        Each species has only one role; we are calculating the dispersion of roles among species with different mean times to extinction. 
        We find that there is a greater variety of roles among species with long times to extinction than short times to extinction.
        The relationship between city coordinates and temperature may be a useful analogy. 
        Temperate cities are very widely dispersed in latitude and longitude, while very hot cities have lower dispersion as they are restricted to the tropics.
        
        
        This difference in dispersion is important because it violates one of the assumptions of the PERMANOVA test.
        This is why we perform supplemental tests and do not place much weight upon the PERMANOVA results.
        We recognise that the concept of role dispersion may be difficult for those not familiar with PERMANOVA tests.
        We therefore have refered only to the violation of assumptions in the Main Text and discuss role dispersion in the Appendix for those who are interested in the details.


    3. Request for explanation of MSEP in main text [done]

        line 172: MSEP is defined in the Appendix but not the main text. I think it would be especially helpful if the text from lines 133-139 describing MSEP and providing a bit more detail on model selection was moved to the main text.
        
        \textbf{R:} In response to Reviewer 2's request for a simplified analysis, we no longer use partial least squares. As such, MSEP is no longer used as a model-selection parameter.


  4. Confusion over de-scaled version of Figure 2 [done]

    line 178: Is the de-scaled version of the model presented in Figure 2 obtained by re-fitting the PLS using the raw values rather than centered and scaled values? If so I think you should explicitly mention this in the methods.   


    \textbf{R:} In response to Reviewer 2's request for a simplified analysis, we no longer use the model described in the old Figure 2. We hope that our current figures are clearer.


  5. Request for discussion of low explanatory power [done]

    Lines 217, 232, 246: Seems like the predictor variables are not explaining much of the variation in mean time to extinction (around 20-25\%), which is not necessarily bad, but what do you think might be missing? Could be worth including in the discussion.

    \textbf{R:}
    The reviewer raises a very good question. In reviewing the manuscript, we now also explore what portion of the variance is explained by trophic level or degree, separately from motif, and find that these species-level descriptors explain up to 0.23\% of the variance. We also discuss other variables which may be important in the discussion.
    
    \begin{quotation}
        Since the trophic level and degree of the species being removed can also affect which species, if any, go secondarily extinct~\citep{Wootton2016a,Dunne2002}, the properties of both disturbed and non-disturbed species affect the ways in which extinctions can cascade through a food web.
        
        The stronger correlation of extinction orders in larger, more-connected networks, as well as the somewhat shorter mean times to extinction in these webs (Fig. S1, \emph{Appendix S1}) may be due to the greater number of short pathways by which an extinction somewhere in the web can affect a focal species. 
		These short pathways are more likely to have strong effects on the population dynamics of species along them than the longer indirect pathways in poorly-connected networks~\citep{Jordan2002,Jordan2006}.
		These stronger effects in turn likely explain why more secondary extinctions in large or highly-connected webs are due to indirect effects rather than direct loss of a prey or predator~\citep{Wootton2016a}. 

    \end{quotation}
    
    We also discuss the importance of link strength, particularly in reference to the omnivory motif.
    
    \begin{quotation}
        This could be related to the distribution of link strengths within a food web.
        In small (3-5 species) model communities and soil food webs, weak interactions only promote stability if omnivory is present~\citep{Neutel2002,Emmerson2004} and the loss of the omnivorous species dramatically reduces stability~\citep{Emmerson2004}.
        In our terms, if the top species in an omnivory motif is lost, the others will be strongly affected.

    \end{quotation}
    
    Finally, in revising the manuscript we now also look at a species \emph{position} within motifs, in addition to simply the motifs it participates in. We find that motif position explains more variance than simply motif role, because some positions are more stabilizing than others.
    
    \begin{quotation}
        Species tended to have longer times to extinction if their roles contained higher frequencies of the bottom position in the apparent competition motif, top position in the direct competition motif, or middle position in the three-species chain.
        Higher frequencies of all other positions, notably all positions in the omnivory motif, were associated with shorter mean persistence times.

    \end{quotation}


  6. Corrections to table references [done]

    line 206: I think you mean to refer to Table S2 in Appendix S3.

    Appendix line 77 is missing a Table reference "Table ??"

    \textbf{R:} Both table references have been corrected. We thank the Reviewer for pointing out the errors.


\clearpage

\Large{\textbf{Reply to Reviewer 2}}

\textbf{R:} Reviewer 2 notes that this is 'a very interesting and original question that is well grounded on earlier work in the literature' and thinks that it 'may help to broaden our theoretical understanding of trophic dynamics'. Their major concerns were that we use only artificial networks and that our methods were too complex. They also give a number of suggestions which have helped to clarify the network. We address all points below.

  \begin{quotation}
  Authors investigate the link between the motif roles of a species, i.e. its frequency of participation to each of the 13 triangular motifs inside a trophic network, and the time before it goes extinct after the primary extinction of another species in the network. Thus, they aim at identifying which motif roles contribute to a species stability or instability given its position in the network. They also try to disentangle the specific effect of the species motif roles from network level effects (richness, connectance) and other species level effects (trophic level, degree).


  This authors treat a very interesting and original question that is well grounded on earlier work in the literature, and which I think may help broaden our theoretical understanding of trophic dynamics. The authors made a noticeable effort in their simulation design to obtain realistic and stable body-mass structured vertebrate food webs.
  % Well that sounds positive!


  (now speaking directly in second person to the authors) However, I don’t understand why you base your study only on artificial networks. Why don’t you do the study on empirical networks instead (or in addition to) the artificial ones? I think the results would be much more convincing with real species interactions and their real body-masses rather than the ones derived from niche model and a few arbitrary heuristics, as you note it yourself (Appendix 1). It seems to me you wouldn’t have to change your methodology to derive the Lotka-Volterra parameters from the network structure. I guess your challenge is that you initially need a network at equilibrium to isolate the effect of forced extinction on the target species extinction time, but you could find an equilibrium state for empirical networks too, or adapt/remove networks which don’t have local stability.
  \end{quotation}


  \textbf{R:} We did not use empirical networks for two reasons, one of which the Reviewer touches upon. First and foremost, there are simply not many empirical networks with attached body mass data for all species. This limits the scope of network size and connectance that can be covered and, especially if the empirical networks were collected by different groups, introduces noise due to different field conditions, sampling strategies, etc. These issues would prevent us from drawing robust conclusions from empirical networks and are the main reason we simulated networks in the first place. While there are also limitations to simulation studies, we feel that the benefits of large sample size, large ranges of size and connectance, and consistent `sampling' methodology are important advantages in our case.
  Second, as the Reviewer notes, our study design requires a web that begins at equilibrium in order to isolate the effect of a species removal. Finding equilibrium states for empirical networks is not straightforward, especially across a large number of networks. On the other hand, removing networks not at equilibrium would diminish an already small sample. While we agree that empirical networks are interesting, the difficulty in finding a large set of sufficiently data-rich, equilibrium networks means that repeating our analyses with empirical networks is beyond the scope of the current study.
  We hope that as sampling methods continuously improve an empirically-based study may become feasible in the future.



  \begin{quotation}
  However, the overall methodology is quite complex, sometimes over-complex in my opinion. Even though I took time to understand it as fully as I could, I couldn’t fully understand the data analysis part, which is problematic as it prevented me to assess the adequation of the conclusions with the evidence. Regarding this statistical methods (e.g. PLS, predictor scaling, see my major comments), not only some points lack clarity, but they often lack a motivation compared to simpler and probably more suited alternatives. My main suggestion is to drastically simplify by removing the PLS, the simple linear regressions (L187 to L193) and you linear mixed models (L194 to 199), and replace all that by 3 multiple linear regressions over all the (6000)networks using time to extinction as the output variable and the following predictors: motif roles (count OR degree standardize OR Z-score), C, R, C:R, degree, STL. I think you would basically get all the answers you are seeking from these 3 models, or I missed something.
  \end{quotation}


  \textbf{R:} We agree that our original methodology was quite complex. 
  Accordingly, we have substantially simplified our approach and hope that it is now clearer.
  However, the proposed multiple linear regression will not suffice to answer our question of interest. We are primarily interested in how motif roles relate to time to extinction and secondarily interested in how motif roles relate to simpler measures (degree and trophic level). With respect to our first question: motif roles, as inherently multivariate descriptors, don't fit well in a linear regression together with univariate predictors. The univariate predictors may have larger effects than any motif individually, but this undervalues the effect of the motif role as a whole. The idea of simplifying a linear regression by removing some elements of a motif role also does not make much sense conceptually.


  Regarding our second question: we could indeed fit and interpret the suggested multiple linear regression. However, this would not tell us how motif roles vary with degree and STL. This is an essential but surprisingly neglected piece of background information that we feel makes an important contribution to efforts to interpret motif roles.
  We appreciate that this second goal was not clear in our previous draft and have revised the manuscript with this in mind.


  \begin{quotation}
  Regarding the form, I think some parts of the methods and results section may be passed to appendix or presented in a much simplified way. I would suggest condensing the three first paragraphs of Results section, which basically say the same things, into one much shorter. One option I would like would be to keep just one type of motif roles in the main manuscript results (the Z-score?), and describe the others in the appendix.
  \end{quotation}


  \textbf{R:}  We thank the Reviewer for the excellent suggestion. We now present the results for species-normalized motifs in detail and have moved the other versions to the Appendix.


  \textbf{Major comments:}

    1. Are perturbation and disturbance interchangeable? [done]

      - L54: Now you use perturbation instead of disturbance: Are you talking about the same concept? You seem to use interchangeably those concepts in the following.

      \textbf{R:} Yes, we are talking about the same concept. We now use `disturbance' exclusively in order to avoid confusion.


    2. Motivation for using stable network unclear

      - L80: Why networks ``networks at stable equilibria''? You don’t justify this methodological choice here, and neither in the methodology section. My guess is that you initially need a network at equilibrium to isolate the effect of the introduced extinction on the targeted species extinction time from a potential effect of an initial disequilibrium state. Whatever is you real reason, it is important to motivate it to justify the extra methodological complexity  required to obtain this equilibrium state.

      \textbf{R:}
      The reviewer is exactly right in their guess. We now state this explicitly.
      
      \begin{quotation}
          We use an equilibrium community to ensure that secondary extinctions are caused by the initial disturbance (i.e. removal of a species) and not due to existing disequilibrium.
      \end{quotation}
      


    3. Specify disturbance earlier

      - (L81) You only specify now (and not even fully) what was meant by disturbance/perturbation to the network: The removal of a species in the network. First, why using the term disturbance which is vague, while you could use extinction instead? Second, why do you chose this type of disturbance? It should be justified regarding the hypothesis you test. Plus, it doesn't appear consistent with the definitions of stability found in the corpus of articles from which you derive your hypotheses on stable motifs: Indeed, both the local stability (May, 1972) and the Quasi Sign stability (Allesina and Pascual, 2008, and later Borrelli, 2015) are related to the stability of populations to small perturbations around the equilibrium state. So why don't you define your disturbance as a small perturbation of species populations in your network instead of an extinction? I quickly read the cited articles and I may have missed some link with your work, but bringing a justification for this choice of disturbance, which determines time to extinction and is thus at the center of your methodology, is critically important.

      \textbf{R:} [done]
      
      The reviewer is correct that stability is a multifaceted concept. We use both 'stability' and 'disturbance' as general terms which relate to other work on, and the importance of, this topic, while clearly defining exactly what we mean by stability and disturbance. Importantly, we are interested in 'stable' motifs, by a number of definitions, and we are here exploring a different dimension of stability. We chose "time to extinction after an initial extinction" as our measure of stability as this gives a clear and dynamic measure of the consequences of a sustained disturbance, such as an extinction. It is also a more realistic representation of the types of disturbances that ecosystems face today than the some of the mathematical definitions of stability such as local stability. We use the word 'disturbance' rather than 'extinction' to avoid confusion between extinction of the initial species and the secondary extinctions it causes. 
      
      We now make this clear in the introduction.
      
      \begin{quotation}
	While 'stability' and 'disturbance' are multifaceted concepts \citep{Donahue2013,Radchuck2019}, we here focus on a dynamic measure of stability to a permanent disturbance --- the time to extinction after a primary extinction. 
      \end{quotation}


    4. Purpose of including trophic level not clear [done]

        -(L77) This is unclear to me why you have to control for trophic level while testing the effect of motif roles on time to extinction? If the purpose is just to test if trophic level, which is indeed a simpler species level measure, is more suited to describe the species stability, then write it more clearly here.
        
        \textbf{R:} As described above, we are not interested in testing whether trophic level is ``more suited'' to describe species stability, we are interested in testing the assertion that motif roles are \emph{also} related to stability.
        As regards trophic level, we are interested in whether any relationship between motif roles and species extinction risk is merely due to underlying relationships between motif roles and trophic level and trophic level and extinction risk.
        Note, however, that motif roles and trophic level are both measurements of a species' place in the network and neither causes the other, such that this relationship is correlative only.
        We therefore want to explicitly identify relationships between a) trophic level and stability and b) trophic level and motifs in order to uncover whether a relationship between c) motifs and stability matches or does not match what we expect base on a) and b).
        We have revised the paragraph to make it clearer why we are including degree and trophic level in our analyses.

        
        Lines 73-86:
      
        \begin{quotation}
            Testing the hypothesis that participation in stable motifs will decrease a species' likelihood of extinction is complicated by the fact that species' motif roles are not independent of other aspects of fine-scale network structure. 
            In particular, a species' degree (number of predators and prey) is likely to strongly affect its motif role.
            The more interaction partners a species has, the more motifs in can participate in (Fig. 1).
            This is likely to lead species with higher degrees to participate in a wider variety of motifs as well as participate more frequently in the four common, stable motifs.
            Likewise, species with different trophic levels tend to have different motif roles~\citep{Cirtwill2018EcolLett}.
            As well as these correlations between simple network measures and motif roles, species with high in-degrees (more prey) or lower trophic levels are also generally less likely to go extinct~\citep{Cirtwill2018FoodWebs}.
            This means that relationships between motif roles and extinction risk could be describing relationships between degree and/or trophic level and both extinction risk and motif roles.
            This possibility should be taken into account when interpreting any such relationship between motif roles and extinction.
        \end{quotation}
      

    5. Missing sentence?

      -l.85 to l.88: I don't get where does that sentence come from. Is that derived from earlier results? Then they should be referred here. Otherwise, this is a result of your study and should be in the results section.

      \textbf{R:} 
     We thank the reviewer for pointing out the gap in our logic here. We now phrase this more clearly as follows.
      
     \begin{quotation}
         Taken together, the results we obtain for these tests show that species are generally consistent in their vulnerability to disturbances...
     \end{quotation}



    6. Vague sentence

      -L127: This last statement is a bit vague and misleading. Your network normalization will tell how the species stability is related with how much it participates to a given motif compared to the participation to this motif over all species of the community. The explanation provided L164 is much better.

      \textbf{R:} We have rephrased this line as quoted below and hope that it is now clearer.

    
        Lines 172-177:
        
        \begin{quotation}
    		The species normalisation allows us to test whether trends in stability with motif participation are due to differences in the total number of motifs a species appears in, while the network normalisation allows us to test whether trends in stability are related to how unusual a species' motif participation is relative to other species in its community, rather than which specific motifs a species participates in.
        \end{quotation}


    7. Justify choice of threshold [Kate to review]

        -L132: Why 1e10-5? Did you found this to be a threshold below which all species tend to vanish? On the contrary, how likely is it that the population grows again after passing under such threshold? Does it exist instead a grounded rule to determine a threshold per species depending on its bodymass and biomass? It would be much better, but I’m not familiar with Lotka-Volterra dynamics.

        \textbf{R:} This is an arbitrary, but small threshold. In Lotka-Volterra simulations, the simulated populations never reach 0, so it is necessary to apply an artificial threshold. Otherwise, a species can sometimes reach a simulated abundance of >>1 individual but then recover to a large population.
        We are not aware of any grounded rule or consensus threshold value to use in such simulations, however. These values are simply chosen to be quite small. For example, comparable simulation studies use thresholds of 1e10-12 (Binzer et al., 2012) and 1e10-20 (Ryser et al., 2019) without an explicit justification of these thresholds in the text. 
        In response to the Reviewer's question and the Editor's suggestion, we have added a brief discussion of what might change with a different choice of threshold in the ``Speculations and Alternative Viewpoints'' section of the discussion. 
        Although establishing a ``best practice'' choice of threshold is beyond the scope of the present manuscript, we hope this acknowledgement of potential threshold effects will ease the Reviewer's mind.
        

    8. Questions about PLS method and study aim [done]

        -L153-156:  I don’t think the PLS method is fitted to the purpose here and neither the simplest method to answer your question on quantifying the individual effect of each motif role/participation on the time to extinction. Regarding parsimony, as you have a ``large'' sample of networks, I don’t see a need for a latent space with reduced dimensionality. Regarding relevance, you wrote earlier (e.g. the previous sentence), and it’s meant by the coef you reconstruct later l.177, that you want here to estimate the effect of each motif role independently (question A), and now you just slide to another question that comes out of the blue: You want to get the effect of participation to a combination of motifs (question B) using PLS. But, if you want to answer question A, which I understood as your real goal, you can use a much simpler method: A multiple linear regression of the ``time to extinction'' (dependent variable) versus the motif roles, and your other variables (explanatory variables).
        
        \textbf{R:}
        The motif role is the 13-dimensional set of participation in all motifs. What we really want to understand is how the 'shape' of this vector relates to time to extinction. 
        The PERMANOVA test tells us whether the motif role as a whole is related to time to extinction, but not which motifs might be particularly important. 
        An equivalent question in community ecology would be `Is diversity different between sites and, if so, what groups drive this difference?` 
        Our main goal is closer to question B, but as question A is simpler we answered it `on the way'.


        As described above, we have simplified our methodology in this revision.
        We have also minimised the discussion of the PERMANOVA in the Main Text.
        We hope that these changes make our focus clearer.
    

    9. Description of STL not clear

      -L186: The description of STL in this sentence is unclear and not consistent with the description in the previous sentence, so you better remove this sentence.

      \textbf{R:} [Done]
      We have now clarified our definition of trophic level as follows.
      
      \begin{quotation}
        Trophic levels, in turn, were calculated based on the food-web structure provided by the niche model. We use the shortest trophic level (STL), i.e. the shortest path between the species and a basal species.
      \end{quotation}
      
      We also simply use "trophic level" where possible to avoid confusion.

    10. Further questions about PLS approach [done]

        -L187: Again, the chosen method doesn't appear fit to the purpose. What you want to do is jointly measure the relative effects of the different predictors, i.e. the motif roles, degree and STL, on the time to extinction, which you can do with a multiple linear regression of the time to extinction (dependent variable Y) versus all predictors at once (explanatory). Actually, a multiple linear regression would answer two of your questions at once, namely the relative effect of each motif role compared to others AND the relative effect of motif roles VS other predictors, and simplify your methodology.
        
        \textbf{R:} Unfortunately, the Reviewer has mis-understood our aim and this comment is rooted in that misunderstanding.
        As stated above, we are not particularly interested in whether degree explains time to extinction better than participation in isolated motifs.
        Rather, we want to know whether motif participation is at all related to extinction risk (this has not yet been shown except in isolated motifs).
        We further want to know how motif participation varies with degree and trophic level, and whether these associations can completely explain any connection between motifs and extinction. 
        Our aim here is to relate a species' motif role more explicitly to other network properties than has been done previously.
        We hope that our simplified methodology and revised text make our goals, and the correspondence between these goals and our models, clearer.


    11. Comment not clear [done]
        
        - L. If your predictors are scaled you should compare directly your the fitted coefficient values between predictors to assess which
        
        \textbf{R:} As part of our methodological changes, we no longer include motifs and degree or trophic level in the same models. We hope that this avoids confusion as we do not have scaled and de-scaled models to compare.


    12. Confusion about de-scaled effects [done]

        -L225: I’m totally lost with this sentence. What do you mean by ``effect'' here? Is it the sum of coefficients of the predictor across PLS axes? Or is it its product with the standard deviation of the predictor across the dataset? Then, you tell that the motif roles (counts) effects appear small because the variables are scaled, but the fitted coefficients should be comparable between variables, no?  Then, you talk about the ``un-scaled effects'' and here I don’t know what you mean and I find myself unable to interpret for comparing STL/degree with motif roles in all PLS models.
        
        \textbf{R:} As part of our methodological changes, we no longer include motifs and degree or trophic level in the same models. We hope that this avoids confusion as we do not have scaled and de-scaled models to compare.


  \textbf{Minor comments:}

    13. Meaning of motif roles not clear in abstract [done]

      L7  I’m unsure about meaning of a species motif roles at this stage. You define it later as the frequency with which it participates in different motifs (L64), but you should introduce it directly here or be more explicit.

      \textbf{R:} We have added a definition of motif roles to the abstract. To avoid confusion we have also retained the definition in the introduction.

      \begin{quotation}
       We test whether a species' time to extinction following a perturbation is related to its participation in stable and unstable motifs and assess how a species' motif role (the frequencies with which the species appears in each motif) co-varies with degree or trophic level.
      \end{quotation}


    14. Potential for larger motifs to affect stability [done]

        L47-51: Another possible interpretation is that higher size motifs (or some whole network architectures), containing the three species motifs mentioned earlier, are determinant for stability. This would mean that the frequencies of the three species motifs themselves are not a driver but is induced by the frequency of higher size motifs. I think that determining if there exist some higher size motifs that explain networks stability better than the three species motifs is an open problem and would deserve discussion.
        
        \textbf{R:} The Reviewer is correct that motifs of different sizes are closely related. However, our understanding is that smaller motifs constrain larger ones. The number of one-way and two-way interactions (two-species motifs) constrain the number of three-way motifs that can occur and whether any of the two-way motifs can occur. Similarly, no four-species motif which includes two predators consuming a common prey can occur if the three-species direct competition motif does not occur. Larger motifs may indeed by a better predictor of stability than three-species motifs, but we are not aware of any studies of these larger motifs which could motivate such a discussion. If the Reviewer can suggest any, we would be happy to include them. In the meantime, we have added a line stating that we focus on three-species motifs since they are the most-commonly studied size of motif with respect to stability.
        
        Lines 35-40:
        
        \begin{quotation}
    	These meso-scale structures, described by frequencies of \emph{motifs} (unique patterns of $n$ interacting species), define the local neighbourhood of a focal species and reflect its direct and close indirect interactions (Fig. 1).
        As three-species motifs are the best studied, particularly with respect to stability~\citep{Stouffer2007,Borrelli2015,Borrelli2015a,Gilling2019}, we also focus on this size of motif.
        \end{quotation}



    15. Confusion about model relating extinction to STL [Done but Alyssa review]

      -L194: In this paragraph, I’m totally lost: I don’t understand the question, nor how the models answer it. Are you trying to test whether the values of degrees/STL may affect the response of time to extinction to motif roles? The sentence must be clarified.

      \textbf{R:} [Done]
      As a result of our revisions, we have simplified this analysis. We now use a simple linear model rather than a mixed effects model. The point of this analysis is to compare the importance of species-level metrics (i.e. trophic level and degree) to motif roles, to learn whether motif roles in fact tell us anything new.
      
      \begin{quotation}
      To interpret the results of the analyses above, especially the LMMs, we must bear in mind the potential for motif participation to reflect differences in degree, trophic level, and global network structure. 
      To establish this context, we therefore fit a series of linear models (LMs) relating degree, trophic level, network size, and connectance to each version of the motif role (\emph{Appendix S6}). 
      \end{quotation}
      


    16. Small language edits [done]

      -L214: ``a species'' is repeated.

        \textbf{R:} Corrected

      -L300: I think ``which'' may be removed from this sentence.

        \textbf{R:} Changed `by which' to `whereby'

      -L306: Except for the omnivory motif, no?

        \textbf{R:} The omnivory motif was notably less stable than the apparent competition, direct competition, and three-species chain motifs (the most stable motifs), despite being much more stable than the 9 other three-species motifs. To avoid any confusion, we now parenthetically specify the three most stable motifs.


    17. Suggestion to do a multiple regression [done]

      - L313: Which is typically what a multiple regression would disentangle…

      \textbf{R:} See response to the Reviewer's opening comments, above.


    18. Request for Lotka-Volterra equations [done]

      Appendix L26: can you be more explicit on how you compute these metabolic rates? Further L46 you also say talk about a consumption rate. In this appendix, can you write the form of the Lotka-Volterra equations as a function of the body masses, the metabolic rates and the consumption rates? It would make the understanding .

      \textbf{R:}
      To clarify this point, we now state the following in the appendix.
      
      \begin{quotation}
      This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details)...... Metabolic and consumptive rates in the Lotka-Volterra model follow a negative quarter power-law relationship with each species' body mass (i.e. proportional to $M^{-0.25}$, where $M$ is the mass of a single, typical adult individual). 
      \end{quotation}
      
      We do not write the full form of the Lotka Volterra equations as they are very long, and they are laid out clearly in \citep{Delmas2017}.
      

    19. Suggestion to move line L111 earlier (to main text?) [done]

      Appendix L111 : This sentence well condensates the question associated with your methodology, but I realise now it didn’t appear as clearly earlier, especially it lacks to the introduction section, you should put it earlier.


      \textbf{R:} The line cited referred to our previous methodology and no longer applies.


    20. Caption of Table S1 unclear. [done]

      - Table S1: The text is unclear: ``Here we show the mean correlation among
      extinction orders across all removed species (R2) and all 100 simulated networks for each combination of S and C...'' -> The mean correlation between extinction orders and what??


      \textbf{R:} As stated in Appendix lines 58-69, this is the mean of correlations of times extinction for each species across all removals. We have slightly revised the caption to make this clearer.


      \begin{quotation}
        Here we show the mean of correlations of extinction orders for a focal species across all removed species ($R^2$)
      \end{quotation}


\end{letter}

\end{document}


